---
title: "Metabolites_mixpopulation"
author: "Sara Nouere"
output: html_document
---
## Introduction
This document analyzes metabolites of the mixed population under different treatments 2023 and 2024. It includes: metabolite differential analysis and volcano plots

**Note:** This R Markdown document is used to generate **Figure S3A-S3B** in the manuscript.

## Load required libraries

```{r,echo=FALSE, message=FALSE, warning=FALSE}
library("dplyr")
library("ggplot2")
library("cowplot")
library("ggpubr")
library("lme4")
library("car")
library("emmeans")
library("readxl")
library("writexl")
library("tidyverse")
library("plotrix") 
library("scales")
library("viridis")
library("ggrepel")
library("tidyr")
library("data.table")
```

## Load and analyze first dataset: 2023

```{r,echo=FALSE, message=FALSE, warning=FALSE}
df <- read.table("data2023_metabolites.txt", sep = "\t", header = TRUE)

data.long <- data %>% gather(metabolite, value, 7:45)
data.filtered <- data.long[!(data.long$value == "NA"), ]
data.filtered$value <- as.numeric(data.filtered$value)
data.filtered$block <- as.character(data.filtered$block)

group.data.long <- group_by(data.filtered, metabolite)
list.mol <- unique(as.character(data.filtered$metabolite))

model_calculator.1 <- sapply(list.mol, function(x) {
  data.s <- group.data.long[group.data.long$metabolite == x, ]
  model <- lmer(value ~ treatment + (1|timepoint) + (1|block), data = data.s)
  aaa <- Anova(model, test = "F")
  aaa$sig <- rep('', length(aaa$`Pr(>F)`))
  makeStars <- function(x) {
    stars <- c("****", "***", "**", "*", "ns")
    vec <- c(0, 0.0001, 0.001, 0.01, 0.05, 1)
    i <- findInterval(x, vec)
    stars[i]
  }
  aaa$sig <- makeStars(aaa$`Pr(>F)`)
  aaa <- aaa["treatment", , drop = FALSE]
  return(aaa)
}, simplify = FALSE, USE.NAMES = TRUE)

res <- do.call(rbind, model_calculator.1)
res <- setDT(res, keep.rownames = TRUE)[]

table.analysis.1 <- data.filtered %>%
  filter(treatment == "control") %>%
  group_by(metabolite) %>%
  summarize(control = mean(value, na.rm = TRUE))

table.analysis.2 <- data.filtered %>%
  filter(treatment == "herbivory") %>%
  group_by(metabolite) %>%
  summarize(herbivory = mean(value, na.rm = TRUE))

res.def <- merge(res, table.analysis.1, by.x = "rn", by.y = "metabolite")
res.def <- merge(res.def, table.analysis.2, by.x = "rn", by.y = "metabolite")
res.def$log2FoldChange <- log2(res.def$herbivory / res.def$control)

de <- res.def
names(de)[names(de) == 'Pr(>F)'] <- 'pvalue'
de$pvalue <- p.adjust(de$pvalue, method = 'fdr')
de$diffexpressed <- "no changes"
de$diffexpressed[de$log2FoldChange > 0.1 & de$pvalue < 0.05] <- "herbivory"
de$diffexpressed[de$log2FoldChange < -0.1 & de$pvalue < 0.05] <- "control"
```

## Load and analyze second dataset

```{r,echo=FALSE, message=FALSE, warning=FALSE}
df <- read.table("data2024_metabolites.txt", sep = "\t", header = TRUE)

data.long.2 <- data.2 %>% gather(metabolite, value, 7:38)
data.filtered.2 <- data.long.2[!(data.long.2$value == "NA"), ]
data.filtered.2$value <- as.numeric(data.filtered.2$value)
data.filtered.2$block <- as.character(data.filtered.2$block)

group.data.long.2 <- group_by(data.filtered.2, metabolite)
list.mol.2 <- unique(as.character(data.filtered.2$metabolite))

model_calculator.2 <- sapply(list.mol.2, function(x) {
  data.s.2 <- group.data.long.2[group.data.long.2$metabolite == x, ]
  model.2 <- lmer(value ~ treatment + (1|timepoint) + (1|block), data = data.s.2)
  aaa.2 <- Anova(model.2, test = "F")
  aaa.2$sig <- rep('', length(aaa.2$`Pr(>F)`))
  makeStars <- function(x) {
    stars <- c("****", "***", "**", "*", "ns")
    vec <- c(0, 0.0001, 0.001, 0.01, 0.05, 1)
    i <- findInterval(x, vec)
    stars[i]
  }
  aaa.2$sig <- makeStars(aaa.2$`Pr(>F)`)
  aaa.2 <- aaa.2["treatment", , drop = FALSE]
  return(aaa.2)
}, simplify = FALSE, USE.NAMES = TRUE)

res.2 <- do.call(rbind, model_calculator.2)
res.2 <- setDT(res.2, keep.rownames = TRUE)[]

table.analysis.1.2 <- data.filtered.2 %>%
  filter(treatment == "control") %>%
  group_by(metabolite) %>%
  summarize(control = mean(value, na.rm = TRUE))

table.analysis.2.2 <- data.filtered.2 %>%
  filter(treatment == "herbivory") %>%
  group_by(metabolite) %>%
  summarize(herbivory = mean(value, na.rm = TRUE))

res.def.2 <- merge(res.2, table.analysis.1.2, by.x = "rn", by.y = "metabolite")
res.def.2 <- merge(res.def.2, table.analysis.2.2, by.x = "rn", by.y = "metabolite")
res.def.2$log2FoldChange <- log2(res.def.2$herbivory / res.def.2$control)

de.2 <- res.def.2
names(de.2)[names(de.2) == 'Pr(>F)'] <- 'pvalue'
de.2$pvalue <- p.adjust(de.2$pvalue, method = 'fdr')
de.2$diffexpressed <- "no changes"
de.2$diffexpressed[de.2$log2FoldChange > 0.1 & de.2$pvalue < 0.05] <- "herbivory"
de.2$diffexpressed[de.2$log2FoldChange < -0.1 & de.2$pvalue < 0.05] <- "control"
```

## Step to sync Y-Axis

```{r,echo=FALSE, message=FALSE, warning=FALSE}
# Get maximum -log10(pvalue) from both datasets
max_y1 <- max(-log10(de$pvalue), na.rm = TRUE)
max_y2 <- max(-log10(de.2$pvalue), na.rm = TRUE)

# Get a common y-axis upper limit
y_max <- ceiling(max(max_y1, max_y2))
```

## Volcano plot for first dataset: 2023

```{r,echo=FALSE, message=FALSE, warning=FALSE}
plot <- ggplot(data = de) +
  theme_bw(base_size = 14) +
  geom_point(aes(x = log2FoldChange, y = -log10(pvalue), colour = diffexpressed), size = 4) +
  geom_text_repel(aes(x = log2FoldChange, y = -log10(pvalue), label = ifelse(diffexpressed != "no changes", rn, "")), max.overlaps = 60) +
  scale_color_manual(values = c("blue", "red", "black")) +
  geom_vline(xintercept = 0, col = "black", linetype = "longdash") +
  geom_hline(yintercept = -log10(0.05), col = "black", linetype = "longdash") +
  scale_y_continuous(limits = c(0, y_max)) +  # <-- 🔥 FIX Y-AXIS
  xlab(expression(paste(Log[2], " Fold Changes"))) +
  ylab(expression(paste(-Log[10], " P"))) +
  theme(
    legend.justification = c(0.99, 0.99),
    legend.position = c(0.99, 0.99),
    panel.grid = element_blank(),
    legend.title = element_blank(),
    legend.background = element_rect(color = NA),
    legend.key = element_rect(color = NA),
    panel.border = element_rect(colour = "black", fill = NA, size = 1)
  )

plot
```

## Volcano plot for second dataset

```{r,echo=FALSE, message=FALSE, warning=FALSE}
plot.2 <- ggplot(data = de.2) +
  theme_bw(base_size = 14) +
  geom_point(aes(x = log2FoldChange, y = -log10(pvalue), colour = diffexpressed), size = 4) +
  geom_text_repel(aes(x = log2FoldChange, y = -log10(pvalue), label = ifelse(diffexpressed != "no changes", rn, "")), max.overlaps = 60) +
  scale_color_manual(values = c("blue", "red", "black")) +
  geom_vline(xintercept = 0, col = "black", linetype = "longdash") +
  geom_hline(yintercept = -log10(0.05), col = "black", linetype = "longdash") +
  scale_y_continuous(limits = c(0, y_max)) +  # <-- 🔥 FIX Y-AXIS
  xlab(expression(paste(Log[2], " Fold Changes"))) +
  ylab(expression(paste(-Log[10], " P"))) +
  theme(
    legend.justification = c(0.99, 0.99),
    legend.position = c(0.99, 0.99),
    panel.grid = element_blank(),
    legend.title = element_blank(),
    legend.background = element_rect(color = NA),
    legend.key = element_rect(color = NA),
    panel.border = element_rect(colour = "black", fill = NA, size = 1)
  )

plot.2
```

## Combine both volcano plots

```{r,echo=FALSE, message=FALSE, warning=FALSE}
combined_plot <- ggarrange(plot, plot.2, labels = c("A", "B"), ncol = 2, nrow = 1)
combined_plot
```