---
title: "Morphology"
author: "Sara Nouere"
output: html_document
---
## Introduction
This document analyzes the morphology: dry weight and surface area in both treatments (control and snail herbivory).

**Note:** This R Markdown document is used to generate **Figure S1A** and **Figure S1B** in the supplementary.

### Figure S1A
## Load required libraries

```{r,echo=FALSE, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(plotrix)   
library(lme4)
library(car)
```

## Load and summarize data

```{r,echo=FALSE, message=FALSE, warning=FALSE}
df <- read.table("morphology_data.txt", sep = "\t", header = T)

mdata.plot <- df %>% group_by(year, timepoint, treatment) %>% dplyr::summarise(area_av = mean(area_fr),
                                                                        area_se = std.error(area_fr),
                                                                        weight_av = mean(biomass_fr),
                                                                        weight_se = std.error(biomass_fr))


transform_x <- function(x) {
  ifelse(x < 18, x,  # Normal spacing before 18
         ifelse(x >= 18 & x <= 48, 18 + (x - 18) / 15,  # Compression for 18-48
                18 + (48 - 18) / 15 + (x - 48) * 1))  # Smooth continuation after 48
}
```

## Plot S2A: biomass per frond

```{r,echo=FALSE, message=FALSE, warning=FALSE}
pxS1A <- ggplot(mdata.plot, aes(x = transform_x(timepoint), y = weight_av, color = treatment)) +
  theme_bw(base_size = 14) +
  geom_pointrange(aes(ymin = weight_av - weight_se, ymax = weight_av + weight_se), position = position_dodge(0.9)) +
  theme(axis.title.x = element_text(color = "black"),
        axis.text.x = element_text(color = "black"),
        axis.text.y = element_text(color = "black"),
        axis.title.y = element_text(color = "black"),
        panel.grid = element_blank(),
        legend.position = "none",
        plot.title = element_text(hjust = 0.5)) +
  scale_color_manual(name = "Legend", values = c("#636363", "#0072B2", "#D55E00"),
                     labels = c("starting", "control", "herbivory"),
                     breaks = c("starting", "control", "herbivory")) +
  scale_x_continuous(
    breaks = transform_x(c(0, 3, 6, 9, 12, 15, 18, 48, 51, 54, 57, 60, 63)),  # Transform the breaks
    labels = c("0", "3", "6", "9", "12", "15", "18", "48", "51", "54", "57", "60", "63")  # Labels remain original
  ) +
  expand_limits(x = c(0, transform_x(65)), y = 0) +
  
  # Grey area for compressed section
  geom_rect(aes(xmin = transform_x(18), xmax = transform_x(48), ymin = -Inf, ymax = Inf), 
            fill = "#B0B0B0", color = "transparent", inherit.aes = FALSE) +
  
  # Add geom_segment layers
  geom_segment(data = mdata.plot,
               aes(x = transform_x(0), xend = transform_x(7),
                   y = weight_av[mdata.plot$treatment == "starting" & mdata.plot$timepoint == 0],
                   yend = weight_av[mdata.plot$treatment == "control" & mdata.plot$timepoint == 7]),
               color = "#0072B2") +
  
  geom_segment(data = mdata.plot,
               aes(x = transform_x(0), xend = transform_x(7),
                   y = weight_av[mdata.plot$treatment == "starting" & mdata.plot$timepoint == 0],
                   yend = weight_av[mdata.plot$treatment == "herbivory" & mdata.plot$timepoint == 7]),
               color = "#D55E00") +

  geom_segment(data = mdata.plot,
               aes(x = transform_x(7) - 0.2, xend = transform_x(11) - 0.2,
                   y = weight_av[treatment == "control" & timepoint == 7],
                   yend = weight_av[treatment == "control" & timepoint == 11]),
               color = "#0072B2") +
  geom_segment(data = mdata.plot,
               aes(x = transform_x(11) - 0.2, xend = transform_x(13) - 0.2,
                   y = weight_av[treatment == "control" & timepoint == 11],
                   yend = weight_av[treatment == "control" & timepoint == 13]),
               color = "#0072B2") +
  
  geom_segment(data = mdata.plot,
               aes(x = transform_x(7) + 0.2, xend = transform_x(11) + 0.2,
                   y = weight_av[treatment == "herbivory" & timepoint == 7],
                   yend = weight_av[treatment == "herbivory" & timepoint == 11]),
               color = "#D55E00") +
  geom_segment(data = mdata.plot,
               aes(x = transform_x(11) + 0.2, xend = transform_x(13) + 0.2,
                   y = weight_av[treatment == "herbivory" & timepoint == 11],
                   yend = weight_av[treatment == "herbivory" & timepoint == 13]),
               color = "#D55E00") +
  
  geom_segment(aes(x = transform_x(13), xend = transform_x(59),
                   y = weight_av[mdata.plot$treatment == "control" & mdata.plot$timepoint == 13],
                   yend = weight_av[mdata.plot$treatment == "control" & mdata.plot$timepoint == 59]),
               color = "#0072B2", linetype = "dashed") +
  
  geom_segment(aes(x = transform_x(13), xend = transform_x(59),
                   y = weight_av[mdata.plot$treatment == "herbivory" & mdata.plot$timepoint == 13],
                   yend = weight_av[mdata.plot$treatment == "herbivory" & mdata.plot$timepoint == 59]),
               color = "#0072B2", linetype = "dashed") +
  
  geom_segment(data = mdata.plot,
               aes(x = transform_x(59) - 0.2, xend = transform_x(61) - 0.2,
                   y = weight_av[treatment == "control" & timepoint == 59],
                   yend = weight_av[treatment == "control" & timepoint == 61]),
               color = "#0072B2") +
  geom_segment(data = mdata.plot,
               aes(x = transform_x(61) - 0.2, xend = transform_x(63) - 0.2,
                   y = weight_av[treatment == "control" & timepoint == 61],
                   yend = weight_av[treatment == "control" & timepoint == 63]),
               color = "#0072B2") +
  
  geom_segment(data = mdata.plot,
               aes(x = transform_x(59) + 0.2, xend = transform_x(61) + 0.2,
                   y = weight_av[treatment == "herbivory" & timepoint == 59],
                   yend = weight_av[treatment == "herbivory" & timepoint == 61]),
               color = "#D55E00") +
  geom_segment(data = mdata.plot,
               aes(x = transform_x(61) + 0.2, xend = transform_x(63) + 0.2,
                   y = weight_av[treatment == "herbivory" & timepoint == 61],
                   yend = weight_av[treatment == "herbivory" & timepoint == 63]),
               color = "#D55E00") +
  
  labs(x = "Time (weeks)",
       y = "Biomass per frond (mg)")

pxS1A
```

## Statistical analysis: biomass per frond

```{r,echo=FALSE, message=FALSE, warning=FALSE}
df1 <- df %>%
  filter(timepoint %in% c(7, 11, 13))

model <- lmer(biomass_fr ~ treatment + (1|timepoint)+ (1|pond), data = df1)
Anova(model,test="F")


df2 <- df %>%
  filter(timepoint %in% c(59, 61, 63))

model <- lmer(biomass_fr ~ treatment + (1|timepoint)+ (1|pond), data = df2)
Anova(model,test="F")
```

### Figure S1B
## Plot S1B: area per frond

```{r,echo=FALSE, message=FALSE, warning=FALSE}
pxS1B <- ggplot(mdata.plot, aes(x = transform_x(timepoint), y = area_av, color = treatment)) +
  theme_bw(base_size = 14) +
  geom_pointrange(aes(ymin = area_av - area_se, ymax = area_av + area_se), position = position_dodge(0.9)) +
  theme(axis.title.x = element_text(color = "black"),
        axis.text.x = element_text(color = "black"),
        axis.text.y = element_text(color = "black"),
        axis.title.y = element_text(color = "black"),
        panel.grid = element_blank(),
        legend.position = "none",
        plot.title = element_text(hjust = 0.5)) +
  scale_color_manual(name = "Legend", values = c("#636363", "#0072B2", "#D55E00"),
                     labels = c("starting", "control", "herbivory"),
                     breaks = c("starting", "control", "herbivory")) +
  scale_x_continuous(
    breaks = transform_x(c(0, 3, 6, 9, 12, 15, 18, 48, 51, 54, 57, 60, 63)),  # Transform the breaks
    labels = c("0", "3", "6", "9", "12", "15", "18", "48", "51", "54", "57", "60", "63")  # Labels remain original
  ) +
  expand_limits(x = c(0, transform_x(65)), y = 0) +
  
  # Grey area for compressed section
  geom_rect(aes(xmin = transform_x(18), xmax = transform_x(48), ymin = -Inf, ymax = Inf), 
            fill = "#B0B0B0", color = "transparent", inherit.aes = FALSE) +
  
  # Add geom_segment layers
  
  geom_segment(aes(x = transform_x(0), xend = transform_x(7),
                   y = area_av[mdata.plot$treatment == "starting" & mdata.plot$timepoint == 0],
                   yend = area_av[mdata.plot$treatment == "control" & mdata.plot$timepoint == 7]),
               color = "#0072B2") +
  
  geom_segment(aes(x = transform_x(0), xend = transform_x(7),
                   y = area_av[mdata.plot$treatment == "starting" & mdata.plot$timepoint == 0],
                   yend = area_av[mdata.plot$treatment == "herbivory" & mdata.plot$timepoint == 7]),
               color = "#D55E00") +
  
  geom_segment(data = mdata.plot,
               aes(x = transform_x(7) - 0.2, xend = transform_x(11) - 0.2,
                   y = area_av[treatment == "control" & timepoint == 7],
                   yend = area_av[treatment == "control" & timepoint == 11]),
               color = "#0072B2") +
  geom_segment(data = mdata.plot,
               aes(x = transform_x(11) - 0.2, xend = transform_x(13) - 0.2,
                   y = area_av[treatment == "control" & timepoint == 11],
                   yend = area_av[treatment == "control" & timepoint == 13]),
               color = "#0072B2") +
  
  geom_segment(data = mdata.plot,
               aes(x = transform_x(7) + 0.2, xend = transform_x(11) + 0.2,
                   y = area_av[treatment == "herbivory" & timepoint == 7],
                   yend = area_av[treatment == "herbivory" & timepoint == 11]),
               color = "#D55E00") +
  geom_segment(data = mdata.plot,
               aes(x = transform_x(11) + 0.2, xend = transform_x(13) + 0.2,
                   y = area_av[treatment == "herbivory" & timepoint == 11],
                   yend = area_av[treatment == "herbivory" & timepoint == 13]),
               color = "#D55E00") +
  
  geom_segment(aes(x = transform_x(13), xend = transform_x(59),
                   y = area_av[mdata.plot$treatment == "control" & mdata.plot$timepoint == 13],
                   yend = area_av[mdata.plot$treatment == "control" & mdata.plot$timepoint == 59]),
               color = "#0072B2", linetype = "dashed") +
  
  geom_segment(aes(x = transform_x(13), xend = transform_x(59),
                   y = area_av[mdata.plot$treatment == "herbivory" & mdata.plot$timepoint == 13],
                   yend = area_av[mdata.plot$treatment == "herbivory" & mdata.plot$timepoint == 59]),
               color = "#D55E00", linetype = "dashed") +
  
  geom_segment(data = mdata.plot,
               aes(x = transform_x(59) - 0.2, xend = transform_x(61) - 0.2,
                   y = area_av[treatment == "control" & timepoint == 59],
                   yend = area_av[treatment == "control" & timepoint == 61]),
               color = "#0072B2") +
  geom_segment(data = mdata.plot,
               aes(x = transform_x(61) - 0.2, xend = transform_x(63) - 0.2,
                   y = area_av[treatment == "control" & timepoint == 61],
                   yend = area_av[treatment == "control" & timepoint == 63]),
               color = "#0072B2") +
  
  geom_segment(data = mdata.plot,
               aes(x = transform_x(59) + 0.2, xend = transform_x(61) + 0.2,
                   y = area_av[treatment == "herbivory" & timepoint == 59],
                   yend = area_av[treatment == "herbivory" & timepoint == 61]),
               color = "#D55E00") +
  geom_segment(data = mdata.plot,
               aes(x = transform_x(61) + 0.2, xend = transform_x(63) + 0.2,
                   y = area_av[treatment == "herbivory" & timepoint == 61],
                   yend = area_av[treatment == "herbivory" & timepoint == 63]),
               color = "#D55E00") +
  
  labs(x = "Time (weeks)",
       y = "Area per frond (mm²)")

pxS1B
```

## Statistical analysis: area per frond

```{r,echo=FALSE, message=FALSE, warning=FALSE}

df1 <- df %>%
  filter(timepoint %in% c(7, 11, 13))

model <- lmer(area_fr ~ treatment + (1|timepoint) + (1|pond), data = df1)
Anova(model,test="F")


df2 <- df %>%
  filter(timepoint %in% c(59, 61, 63))

model <- lmer(area_fr ~ treatment + (1|timepoint) + (1|pond), data = df2)
Anova(model,test="F")
```

